Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-05-06 Manuscript/manuscript.Rmd
methods edits (computational framework for boundary prediction section)

## Computational framework for TAD boundary prediction

@@

To create shifted binning, the first chromosome-specific bin was set at half of the resolution *r*, and continued in congruent intervals of length *r* until the end of the chromosome (*mod r* + *r/2*). The shifted genomic bins, referred hereafted as bins for simplicity, were then defined as TAD boundary regions (*Y = 1*) if they contained a TAD boundary, and non-TAD boundary regions (*Y = 0*) otherwise, thus establishing the binary response vector (**Y**) used for classification (Figure 1).

Cell type-specific genomic annotations were used to build the predictor space for our TAD boundary prediction framework. Bins were annotated by the number of overlaps with an annotation (*Overlap Count (OC)*), the percent of overlap between the bin and the total width of genomic annotation regions overlapping it (*Overlap Percent (OP)*), the average [??? Is that correct? Or, max?] signal strength of the corresponding annotation (*Peak Signal Strength*), and the distance in bases from the center of the bin to the center of the nearest genomic annotation region (*Distance*). A log2-transformation of distance was used to account for the skewness present in the distributions of genomic distances (Supplementary Figure 1).

Chromosome-specific datasets used for modeling were composed of two components: the response vector $Y_N$, an indicator variable separating between TAD and non-TAD boundary regions, and the predictor-type specific data matrix, $X = X_1, X_2, ..., X_p$ [??? Need to incorporate N, use proper math notation, maybe $\in$], where $n \in N$ is the number of bins (different for each chromosome) and $p$ is the number of genomic annotations (depends on the type of annotartions). Chromosome-specific dataseta were concatenated to form a whole genome dataset. The dataset was then split into a 7:3 ratio for training and testing, respectively.

>>>>

## Computational framework for boundary prediction

@@

To create shifted binning, the first chromosome-specific bin was set at half of the resolution *r*, and continued in congruent intervals of length *r* until the end of the chromosome (*mod r* + *r/2*). The shifted genomic bins, referred hereafted as bins for simplicity, were then defined as boundary regions (*Y = 1*) if they contained a TAD (or loop) boundary, and non-boundary regions (*Y = 0*) otherwise, thus establishing the binary response vector (**Y**) used for classification (Figure 1).

Cell type-specific genomic annotations were used to build the predictor space for our TAD boundary prediction framework. Bins were annotated by the number of overlaps with an annotation (*Overlap Count (OC)*), the percent of overlap between the bin and the total width of genomic annotation regions overlapping it (*Overlap Percent (OP)*), the average [??? Is that correct? Or, max?; !!!yes average] signal strength of the corresponding annotation (*Peak Signal Strength*), and the distance in bases from the center of the bin to the center of the nearest genomic annotation region (*Distance*). A log2-transformation of distance was used to account for the skewness present in the distributions of genomic distances (Supplementary Figure 1). Note that model built using a *Peak Signal Strength* feature space where only composed of histone modifications and transcription factor binding sites. That is, they did not include BroadHMM chromatin segmentations.

Chromosome-specific datasets used for modeling were composed of two components: the response vector $Y_{i}$, an indicator variable separating between boundary and non-boundary regions, and the predictor-type specific data matrix, $\textbf{X}_{ij}$, where $i \in \{ 1,2, \cdots,N \}$ ($N$ being the number of chromosome-specific bins) and $j \in \{ 1,2,\cdots,p \}$ ($p$ being the number of genomic annotations). <!--Chromosome-specific datasets were concatenated to form a whole genome dataset. The dataset was then split into a 7:3 ratio for training and testing, respectively.[!!!only comes into play with preciseTAD]-->
